COVID Shutdown Regulations Negatively Impact Employment and Sales of Restaurants in Ontario

### 2021-02-27 outputs/paper/paper.Rmd
Proofreading

---
title: "COVID Shutdown Regulations Negatively Impact Employment and Sales of Restaurants in Ontario"
author: "Hong Shi, Hong Pan, Yixin Guan, Babak Mokri"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Due to the world-wide COVID-19 pandemic, shutdown restrictions on businesses and facilities have limited economic activities into a recession. This paper examines the shutdown effect on restaurants by reopening a city in Ontario (Peterborough) and comparing its restaurant operating situation with that of another city (Brantford) after three months reopening. The survey result indicates that shutdown restrictions have negative impacts on restaurant employment, sales and operating confidence. However, the higher increased COVID cases in the reopened city sugguests the Ontario government to mindfully plan for futher policy decisions."
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
toc: FALSE

@@

&nbsp; &nbsp; To help recover from the current recession, we at Petit Poll have partnered with the Ontario government to examine the effects and scope of COVID shutdown on restaurants. We conducted an experiment by selecting two single tier cities, namely Peterborough and Brantford, ^[Ontario's Municipal Act, 2001 defines a single-tier municipality as "a municipality, other than an upper-tier municipality, that does not form part of an upper-tier municipality for municipal purposes"] with similar restaurant operating situations in Ontario and investigated how reopening the dine-in services for restaurants in Peterborough has prevailed over the state of continued shutdown in Brantford through surveys. The results showed that continuing shutdown has negatively impact the employment, sales volume and operating confidence on restaurant businesses in Brantford. In addition, the governmental support programs contributed substantially to restaurant businesses in both reopened and shutdown cities. However, the higher increased COVID cases in Peterborough demonstrates the drawback of the reopening, and the Ontario government should hold a clear measure between public health and economic recovery before further policy implementations.

&nbsp; &nbsp; We organize the rest of paper as follows: In Data section, we first discuss interventions of the experiment. Next, we introduce sampling methods and survey methodologies to collect the data. And then, we discuss main findings from survey results. In Discussion section, we would elaborate on concerns about the intervention on city selections, sampling methods and survey methodologies. And we also propose our thoughts on further policy implications. In Appendix, we include screenshots of survey contents.^[ Code and data are available at the GitHub repo: https://github.com/honn-ishinn/restaurants_covid_shutdown"]

@@

&nbsp; &nbsp; Partnered with the Ontario government, we want to examine the shutdown effects on restaurant businesses in Ontario. The plan of the experimental design was to reopen all restaurants (i.e. enable the dine-in option) in particular Ontario regions, while enforcing current provincewide shutdown restrictions (i.e. disable the dine-in option) on restaurants in the rest of Ontario regions. And after the three-month reopening, we compare restaurant operating situations in reopened regions with those in shutdown regions through data collected from online surveys of restaurant owners.

&nbsp; &nbsp; However, due to current COVID pandemic, any loosening of shutdown restrictions increases chances of COVID exposures, resulting in challenges on public health and healthcare of reopened regions. It is highly risky and hasty to reopen a large area of Ontario regions for the sole purpose of our experiment. Therefore, we mindfully chose Peterborough, a relative small size Single-tier city 125 kilometers northeast of Toronto, as the only region to reopen in Ontario and then compared its restaurant operating situations with those in Brantford, another Single-tier city 105 kilometers southwest of Toronto with similar population. Since demographics, land areas and more importantly, restaurant operating situations between these two cities are quite similar with each other^[Approaches to select these two cities are further introduced in the discussion section], we consider these two cities as valid comparison cities that could yield an accurate estimate of restaurant businesses shutdowns. In our experiment, we assigned restaurants in Peterborough as the treatment group and those in Brantford as the control group. We also ensured restaurants within these two cities to have equal opportunities receiving government support (e.g. subsidies, grants, etc.) to estimate the true impact of shutdown effects on restaurant businesses after three-month reopening in Peterborough.

&nbsp; &nbsp; In September, 2020, we at Petit Poll conducted the first survey on restaurants in Peterborough and Brantford to examine their business conditions before the experiment. Starting from October, 2020, the Ontario government reopened Peterborough while keeping shutdown restrictions on the rest of Ontario regions. After a three month period in January 2021, we at Petit Poll started to conduct survey on restaurants in Peterborough and Brantford about their business performance from October to December. Survey results were released in February 2021.

@@

&nbsp; &nbsp; As we would like to understand the effect of COVID shutdowns on restaurant businesses in Ontario, we considered all restaurants in Ontario as our target population. According to our intervention, the frame population are all restaurants in Peterborough and Brantford. We used the following approaches to obtain the sampled population as below:

1. Scrape postal codes of all restaurants located within the circle of the Brantford City Centre to a 40 km radius on Yelp. And then apply the same method for Peterborough.^[Land areas of both Brantford and Peterborough are less than 100 kilometer squares, so searching restaurants 40 km within the city center almost ensures to include all restaurants in Peterborough and Brantford]

2. Acquire a list of postal codes in Brantford CMA and Peterborough CMA from the Postal Code Conversion File (PCCF) retrieved from the University of Toronto Library[@cite_uoftlibrary].

3. Compare these two lists and eliminate restaurants with postal codes outside of Peterborough CMA and Brantford CMA to get the list of all restaurants in Brantford and Peterborough as the frame population.

4. Randomly select 100 restaurants from Brantford and 100 restaurants from Peterborough to form the sampled population.

&nbsp; &nbsp; The first survey was conducted before the decision on intervention to ensure our assumptions about restaurant similarities between Peterborough and Brantford. We asked selected respondents about the restaurants' employment, average monthly sales, nature and category of restaurant, restaurant capacity to examine restaurant similarities between these two cities.

@@

&nbsp; &nbsp; We used telephone interviews and online questionnaire methods to administer the survey. The telephone interview served as the primary channel and an online survey option was provided when respondents could not or prefer not to be interviewed over the phone. As shown in the above table (Table \@ref(tab:presamplingtable)), our sampled population included 50 valid responses in Brantford and 55 in Peterborough. And among which, 7 were collected via the online survey in Brantford and 8 were collected via the online survey in Peterborough. Online surveys with no or incomplete answers were treated as nonresponses and were not included in the dataset for further analysis^[Exception for Question 6 of the first survey, which respondents could skip the question]. The survey was administered in following steps:

@@

4. To ensure the quality and accuracy of data collection, survey staffs would ask if the respondent at the restaurant worked at a management-level position in the restaurant or was familiar with the restaurant’s sales performance. If the respondent answered no, the survey staff would ask the caller to pass the call to the relevant staff in the restaurant. If that staff was not available, the survey staff will schedule another call or provide the online questionnaire based on the preference identified by the respondent. Respondents who were interested in answering questions via online questionnaires could leave their email addresses and have two weeks to complete survey questions.

@@

kable( f_survey[,c(4,1,2,3)],caption = "Summary Statistics of the First Survey",
       booktabs = TRUE,
       valign='t') 
```
&nbsp; &nbsp; From the summary statistics of the first survey as shown above (Table \@ref(tab:survey1), values inside the bracket represent percentage), we find that the distribution of category, nature, number of employees, average monthly sales and capacity of restaurants is closely identical between Peterborough and Brantford. These results confirm our initial assumptions that restaurants in these two cities are similar enough to implement our intervention of reopening the Peterborough.

@@

&nbsp; &nbsp; The second survey examined two parts of our interests: (1) Effects of the intervention on restaurant business conditions. (2) Effects of the government support during the intervention. Respondents answered survey questions based on business situations during the intervention period. Several key findings are discussed in the following:

@@

&nbsp; &nbsp; Compared to Peterborough, restaurants in Brantford have much higher number of layoffs (Figure \@ref(fig:question3)), suggesting that the continuing shutdown greatly affected the employment situation in Brantford. In Peterborough, there are considerable amount of restaurants with no layoffs, implying that restaurants would require certain number of staffs to operate the dine-in service after reopening.

@@

q4_table <- data.frame(table(simulated_dataset$q4_total_sales,simulated_dataset$q1_city))
data.frame(table(simulated_dataset$q4_total_sales,simulated_dataset$q1_city))%>%
  mutate(Var1 = fct_relevel(Var1,"Prefer not to say", "Decreased more than 10%","Decreased 5% ~ 10%","Roughly unchanged","Increased 5% ~ 10%","Increased more than 10%")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
  geom_bar(stat = "identity",width = 0.8, position = position_dodge()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Total Sale Change")+
  ggtitle("Total Sales Volume Change Oct to Dec vs July to Sept in 2020")+
  guides(fill=guide_legend(title="City"))+
  coord_flip()
```
&nbsp; &nbsp; From Figure \@ref(fig:q4), we notice a considerable increase in sales in Peterborough during the reopen period. This indicated that if restaurants were open, people would had used their dine-in services, regardless of potential pandemic risks.

&nbsp; &nbsp; In contrast, continuing shutdown restrictions during October to December on Brantford resulted in a slight decrease in sales volume for these restaurants. We propose possible reasons for the decrease in restaurant sales: (1) Residents in Brantford changed their mealing behaviors to cook more by themselves, instead of relying on meals from restaurants. (2) The continuing shutdown resulted in closures of other businesses, so these business staffs were not contributing for restaurant sales anymore. (3) There exists some seasonal fluctuations on food service industries.

@@

&nbsp; &nbsp; Increased costs of restaurants in Peterborough, as shown in Figure \@ref(fig:q5), demonstrated more activities in their restaurants. This change admits that restaurants would have preserved more of their staff and consumed more utility or housekeeping materials to maintain their dine-in area. Although increased operating cost would usually impact the business unit negatively, from a broader view, these costs generate economic flow into local society and are vital for other business sectors to survive. So we consider the increased cost in restaurants during the reopening period as a good signal of economic wellbeing.

&nbsp; &nbsp; Figure \@ref(fig:q8) shown below illustrates respondents' opinions towards the expected survivability of their businesses. It is very unfortunate that restaurants in Brantford are not as positive to continue their businesses for longer times, indicating a stressful operating situation during this continuing shutdown period. While on the other hand, restaurants in Peterborough generally expect to survive longer, showing stronger confidence to continue their businesses after reopening.

@@

&nbsp; &nbsp; In addition, survey results about expected recovery time back to pre-COVID situation (Figure \@ref(fig:q10)) show that even if the pandemic ends immediately, most restaurants in Brantford, especially quick-service restaurants, hold negative views to recover back to pre-COVID operating situation in a short period. For restaurants in Peterborough, despite some suspicions to recoup with the pre-COVID situation in more than a year, most of them expected to recovery in less than a year. Therefore, we believe that the continuing shutdown in Brantford have harmed expectations of future operation possibilities in Brantford, while the reopening in Peterborough could build confidence to alleviate current operation difficulties.

@@

&nbsp; &nbsp; As agents to gather relevant feedbacks for the Ontario Government, we asked respondents' opinions towards the helpfulness of government financial support programs. Survey results (Table \@ref(tab:q11)) show that both cities generally treat support programs as helpful, especially in Brantford where the shutdown restriction was still implementing. From previous gloom results on restaurant business conditions, it is not a surprise that restaurants in Brantford require more government financial supports.

@@

# Question 11 To what extent do you feel the government financial support programs helped your business survive? (1 being terrible, 5 being excellent.)
simulated_dataset %>%
  tabyl(q1_city, q11_rate_support)%>%
  adorn_totals("row", name = "Average")%>%
  adorn_percentages("row")%>%
  adorn_pct_formatting(digits=2)%>%
  adorn_ns()%>%
  adorn_title("combined",row_name = "City", col_name = "Rating")%>%
  rename(
    `Terrible` = `1`,
    `Poor` = `2`,
    `Fair` = `3`,
    `Good` = `4`,
    `Excellent` = `5`
  ) %>% 
  kable(align = 'ccccccc', caption = 'Helpfulness of Government Financial Support Programs on Restaurants') %>% 
  kable_styling(latex_options = "hold_position")
```

As shown in Figure \@ref(fig:q14), restaurants in Peterborough have applied more business loans than those in Brantford during the intervention. Since loans have to be paid back, businesses at low operating confidence are less likely to apply for loans. With increased business opportunities after reopening, restaurant owners were more likely to invest for faster recovery, and this corresponds with our previous findings of longer expected survival time (Figure \@ref(fig:q8)) and expected recovery time (Figure \@ref(fig:q10)) in Peterborough restaurants as mentioned before.

@@

In contrast, restaurants in Brantford generally receives more grants than those in Peterborough(Figure \@ref(fig:q15)). Since grants are non-repayable funds and need certain conditions to apply, restaurants in Peterborough with better business conditions may not be eligible to apply for grants. And restaurants that suffer more in Brantford could receive more grants during the shutdown.

@@

To obtain an accurate estimate of the shutdown effect on restaurant businesses, average characteristics of restaurants in the reopening city and those in the shutdown city need to be identical in the absence of the reopen intervention under current COVID outbreak. Since businesses in the food service industry are highly labor intensive [@cite_labor], the `total employment` on food service between selected cities is preferred to be equal as the identical characteristic. Besides, `population density` ( `Total Population`/ `Total Land Area in squared kilometers`) also has significant organizational life-cycle effects[@cite_density]. At different population density levels, competitions for food services varies, leading to different restaurants mortality rates even without the reopen intervention. As a result, total employment on food service and population density serve as primary restaurant operation characteristics for the selection of treatment city and control city.

@@

We used the employment by industry data and 2016 Census data available at Statistics Canada to identify cities with similar total employment on food service and population density. The above point graph(Figure \@ref(fig:selectplot)) indicates that Brantford and Peterborough have similar characteristics affecting restaurant operations. In addition, the coefficient of the graph $$\frac{Population\:Density}{Total \:Employment}=\frac{Total\:Population/Total\:Land\:Area}{Total\:Employment}$$, which could be interpreted as the number of people served by a restaurant employee per square kilometers, helps to identify the restaurant operating situation in designated regions. Similar coefficients between Brantford and Peterborough (0.306 vs 0.308) also suggest close restaurant characteristics. Besides, there is neither significant demographics nor land areas difference between these two cities, so we consider them as a valid comparison group and reopened Peterborough as the treatment group while maintaining Brantford shutdown as the control group.

@@

* The employment data at Statistics Canada uses the NAICS (North American Industry Classification System) standard and assigns employment services to Accommodation and Food Service category. So there is less employment in food services. We took the convention that the ratio between the accommodation employment and the food service employment between Peterborough and Brantford is the same.

* Since the distance between Brantford and Peterborough is more than 200 kilometers apart, residents in Brantford are unlikely to travel such long distance solely for a dine-in meal. So the treatment of reopening Peterborough would not affect the mealing behavior in Brantford. However, residents in shutdown regions near Peterborough might go for dine-in meals, potentially increasing sales in Peterborough restaurants. The actual shutdown effect on restaurants could be overestimated due to possible increase in sales from nearby shutdown regions' residents.

* The mindful comparison between cities' similarities on employment, population density, etc. ensures the internal validity of our experiment so that we could accurately estimate the true impact of shutdown restrictions on restaurant. However, our shutdown effect evaluations may not be generalized to the entire population of interest, i.e. all restaurants in Ontario such as township regions or cosmopolitan city Toronto due to different levels of competitions, operating costs, etc.

@@

* Since results of the first survey before implementing our experiment suggest apparent similarities between restaurants in Peterborough and those in Brantford, we assume surveyed restaurants to have similar qualities. However, attributes of restaurants (capacity, employment, service type, etc) were not controlled in our analysis. In other words, we believe the response rate within the different restaurant subgroups to be the same. Variations in the actual response between restaurants in two cities are unknown and may impact the internal validity of results.

* Even though the size our sampled population is somehow small (100 restaurants each in Peterborough and Brantford), it has already captured more than half of our frame population (Therer are no more than 200 restaurants in each city). So we believe our sampling population could represent general characteristics of our frame population.

* The sample suffered from survivorship bias as we only surveyed restaurants that were still in operations. Restaurants that lost their businesses during the pandemic before October in 2020 could not be captured with our intervention started in October, 2020, while these restaurants would be most impacted by the COVID shutdowns. Another aspect of the survivorship bias involves the use of the same frame population for the first survey and the second survey. The three-month intervention would also affect business operating situations of restaurants. And the decreased response rate of the second survey might also imply the survivorship bias of our experiment.

* The sampling methods bore self-selection bias. From a psychological standpoint, respondents may have a stronger desire to respond the survey if they have a strong opinion to express about their situations. Because of increasing of sales and  growing operation confidence during the intervention, respondents in Peterborough restaurants might be more willing to participate to our survey, leading to the higher response rate in Peterborough.

@@

* Compared to the average survey response rate around `33%` [@cite_responserate], the response rates of our surveys are relatively high (more than `50%` in the first survey and `40%` in the second survey). We credit our high response rates to efforts made in the following: (1) We spent a great amount of our budget to hire and train survey staffs to better approach our respondents. (2) We offered respondents chances to withdraw gift cards as incentives to participate our surveys. (3) We did not ask exact value of variables of our interest (e.g. Sales). Instead, we offer respondents several options to choose (e.g. `$0 ~ $10,000` ) so that respondents are more comfortable to participate in our survey.

>>>>

---
title: "COVID Shutdown Regulations Negatively Impact Employment and Sales of Restaurants in Ontario"
author: "Hong Shi, Hong Pan, Yixin Guan, Babak Mokri"
date: "February 26 2021"
abstract: "Due to the world-wide COVID-19 pandemic, shutdown restrictions on businesses and facilities have limited economic activities. This paper examines the shutdown effect on restaurants by reopening a city in Ontario (Peterborough) and comparing its restaurant operating situation with that of another city (Brantford) after three months. The survey results indicate that shutdown restrictions have negative impacts on restaurant employment, sales and operating confidence. However, the more drastic increase in COVID-19 positive cases in the reopened city suggests the Ontario government to be conservative for further policy decisions."
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
toc: FALSE

@@

&nbsp; &nbsp; To help recover from the current recession, we at Petit Poll have partnered with the Ontario government to examine the effects and scope of COVID shutdown on restaurants. We conducted an experiment by selecting two single tier cities, namely Peterborough and Brantford, ^[Ontario's Municipal Act, 2001 defines a single-tier municipality as "a municipality, other than an upper-tier municipality, that does not form part of an upper-tier municipality for municipal purposes"] with similar restaurant operating situations in Ontario and investigated how reopening the dine-in services for restaurants in Peterborough has prevailed over the state of continued shutdown in Brantford through surveys. The results showed that the reopening policy has positively impacted the employment, sales volume and operating confidence on restaurant businesses in Peterborough. In addition, the governmental support programs contributed substantially to restaurant businesses in both reopened and shutdown cities. However, the more drastic increase in the COVID-19 positive cases in Peterborough demonstrated the public health concerns of the reopening, and the Ontario government should strike a balance between public health and economic recovery in further policy decisions.

&nbsp; &nbsp; We organize the rest of paper as follows: In Data section, we first discuss interventions of the experiment. Next, we introduce sampling methods and survey methodologies to collect the data. And then, we discuss main findings from survey results. In Discussion section, we elaborate on concerns about the intervention on city selections, sampling methods and survey methodologies. And we also propose our thoughts on further policy implications. In Appendix, we include screenshots of survey contents.^[ Code and data are available at the GitHub repo: https://github.com/honn-ishinn/restaurants_covid_shutdown"]

@@

&nbsp; &nbsp; We want to examine the shutdown effects on restaurant businesses in Ontario. The plan of the experimental design was to reopen all restaurants (i.e., enable the dine-in option) in particular Ontario regions, while enforcing current provincewide shutdown restrictions (i.e., disable the dine-in option while enable the take-out option) on restaurants in the rest of Ontario regions. And after the three-month reopening, we compare restaurant operating situations in reopened regions with those in shutdown regions through data collected from online surveys of restaurant owners.

&nbsp; &nbsp; However, due to the current COVID-19 pandemic, any loosening of shutdown restrictions increases chances of COVID exposures, resulting in challenges on public health and safety of reopened regions. It is highly risky and hasty to reopen a large area of Ontario regions for the sole purpose of our experiment. Therefore, we mindfully chose Peterborough, a relatively small size Single-tier city 125 kilometers northeast of Toronto, as the only region to reopen in Ontario. We compared Peterborough's restaurants operating situations with those in Brantford, another Single-tier city 105 kilometers southwest of Toronto with similar population. Since demographics, land areas and more importantly, restaurant operating situations between these two cities are quite similar with each other^[Approaches to select these two cities are further introduced in the discussion section], we consider these two cities as valid comparison cities that could yield an accurate estimate of restaurant businesses shutdowns. In our experiment, we assigned restaurants in Peterborough as the treatment group and those in Brantford as the control group. We also ensured restaurants within these two cities to have equal eligibilities to receive government supports with certain conditions(e.g. subsidies, grants, etc.) to estimate the real impact of shutdown effects on restaurant businesses after three-month reopening in Peterborough.

&nbsp; &nbsp; In September 2020, Petit Poll conducted the first survey on restaurants in Peterborough and Brantford to examine their business conditions before the experiment. Starting from October 2020, the Ontario government reopened Peterborough while keeping shutdown restrictions on the rest of Ontario regions. After a three-month period in January 2021, Petit Poll started to conduct survey on restaurants in Peterborough and Brantford about their business performance from October to December 2020. Survey results were released in February 2021.

@@

&nbsp; &nbsp; As we would like to understand the effect of COVID shutdown on restaurant businesses in Ontario, we considered all restaurants in Ontario as our target population. The frame population are all restaurants in Peterborough and Brantford. We used the following approaches to obtain the sampled population as below:

1. Scraped postal codes of all restaurants located within the circle of the Brantford City Centre to a 40 km radius on Yelp. And then applied the same method for Peterborough.^[Land areas of both Brantford and Peterborough are less than 100 kilometer squares, so searching restaurants 40 km within the city center almost ensures to include all restaurants in Peterborough and Brantford]

2. Acquired a list of postal codes in Brantford and Peterborough from the Postal Code Conversion File (PCCF) retrieved from the University of Toronto Library[@cite_uoftlibrary].

3. Compared these two lists and eliminate restaurants with postal codes outside of Peterborough and Brantford to get the list of all restaurants in Brantford and Peterborough as the frame population.

4. Randomly selected 100 restaurants from Brantford and 100 restaurants from Peterborough to form the sampled population.

&nbsp; &nbsp; The first survey was conducted before the decision on intervention to ensure our assumptions about restaurant similarities between Peterborough and Brantford. We asked respondents about the restaurants' employment, average monthly sales, nature and category of restaurant, restaurant capacity to examine restaurant similarities between these two cities.

@@

&nbsp; &nbsp; We used telephone interviews and online questionnaire methods to administer the survey. The telephone interview served as the primary channel and an online survey option was provided when respondents could not or prefer not to be interviewed over the phone. As shown in the above table (Table \@ref(tab:presamplingtable)), our sampled population included 50 valid responses in Brantford and 55 in Peterborough.Among which, 7 were collected via online surveys in Brantford and 8 in Peterborough. Online surveys with no or incomplete answers were treated as nonresponses and were not included in the dataset for further analysis^[Exception for Question 6 of the first survey, which respondents could skip the question]. The survey was administered in following steps:

@@

4. To ensure the quality and accuracy of data collection, survey staffs would ask if the respondent at the restaurant worked at a management-level position in the restaurant or was familiar with the restaurant’s sales performance. If the respondent answered no, the survey staff would ask the caller to pass the call to the relevant staff in the restaurant. If that staff was not available, the survey staff would schedule another call or provide the online questionnaire based on the preference identified by the respondent. Respondents who were interested in answering questions via online questionnaires could leave their email addresses and have two weeks to complete survey questions.

@@

kable( f_survey[,c(4,1,2,3)],caption = "Summary Statistics of the First Survey",
       booktabs = TRUE,
       valign='t') 
```
&nbsp; &nbsp; From the summary statistics of the first survey as shown above (Table \@ref(tab:survey1), values inside the bracket represent percentage), we found that the distribution of category, nature, number of employees, average monthly sales and capacity of restaurants was similar between Peterborough and Brantford. These results confirm our initial assumptions that restaurants in these two cities are similar enough to implement our intervention of reopening Peterborough.

@@

&nbsp; &nbsp; The second survey examined two parts of our interests: (1) Effects of the intervention on restaurant business conditions. (2) Effects of the the intervention on restaurants' dependency on government supports. Respondents answered survey questions based on business situations during the intervention period. Several key findings are discussed in the following:

@@

&nbsp; &nbsp; Compared to Peterborough, restaurants in Brantford had much higher number of layoffs (Figure \@ref(fig:question3)), suggesting that the continuing shutdown greatly affected the employment situation in Brantford. In Peterborough, there were considerable number of restaurants with no layoffs, possibly because restaurants required staffs to operate the dine-in service after reopening.

@@

q4_table <- data.frame(table(simulated_dataset$q4_total_sales,simulated_dataset$q1_city))
data.frame(table(simulated_dataset$q4_total_sales,simulated_dataset$q1_city))%>%
  mutate(Var1 = fct_relevel(Var1,"Prefer not to say", "Decreased more than 10%","Decreased 5% ~ 10%","Roughly unchanged","Increased 5% ~ 10%","Increased more than 10%")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
  geom_bar(stat = "identity",width = 0.8, position = position_dodge()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Total Sale Change")+
  ggtitle("Total Sales Volume Change Oct to Dec vs July to Sept in 2020")+
  guides(fill=guide_legend(title="City"))+
  coord_flip()
```
&nbsp; &nbsp; From Figure \@ref(fig:q4), we noticed a considerable increase in sales in Peterborough during the reopen period. This indicated that when restaurants opened, some people would have used their dine-in services, regardless of potential pandemic risks.

&nbsp; &nbsp; In contrast, continued shutdown restrictions during October to December on Brantford resulted in a slight decrease in sales volume for these restaurants. Possible reasons for the decrease in restaurant sales included: (1) Residents in Brantford changed their preferences to cook more by themselves, instead of relying on meals from restaurants. (2) The continued shutdown resulted in closures of other businesses, therefore, minimizing their customer volume. (3) There exists some seasonal fluctuations on food service industries. People were more likely to stay at home when the weather was cold.

@@

&nbsp; &nbsp; Increased costs of restaurants in Peterborough, as shown in Figure \@ref(fig:q5), demonstrated more activities in their restaurants.We expect the cost increase would likely be variable cost. This change suggested that restaurants would have spent more in hiring staff, utilities or housekeeping materials to maintain their dine-in area. Although increased operating cost would usually impact the business unit negatively, from a broader view, these costs could generate economic flow into local society and are vital for other business sectors to survive. Therefore, we consider the increased cost in restaurants during the reopening period as a good signal of economic wellbeing.

&nbsp; &nbsp; Figure \@ref(fig:q8) shown below illustrates respondents' opinions towards the expected survivability of their businesses. Restaurants in Brantford were not as positive to continue their businesses for longer times, indicating a stressful operating situation during this continuing shutdown period. While on the other hand, restaurants in Peterborough generally expected to survive longer, showing stronger confidence to continue their businesses after the pandemic.

@@

&nbsp; &nbsp; In addition, survey results about expected recovery time back to pre-COVID situation (Figure \@ref(fig:q10)) showed that even if the pandemic ends immediately, most restaurants in Brantford, especially quick-service restaurants, would hold negative views to recover back to pre-COVID operating situation in a short period. For restaurants in Peterborough, despite some suspicions to recoup with the pre-COVID situation in more than a year, most of them expected to recovery in less than a year. Therefore, we believe that the continuued shutdown in Brantford had harmed the business confidence of Brantford restaurants, while the reopening in Peterborough built confidence to alleviate current operation difficulties.

@@

&nbsp; &nbsp; As agents to gather relevant feedbacks for the Ontario Government, we asked respondents' opinions towards the helpfulness of government financial support programs. Survey results (Table \@ref(tab:q11)) showed that both cities generally considered support programs helpful, especially in Brantford where the shutdown restriction was still being  implemented. From previous gloom results on restaurants' business conditions, it is not a surprise that restaurants in Brantford required more government financial supports.

@@

# Question 11 To what extent do you feel the government financial support programs helped your business survive? (1 being terrible, 5 being excellent.)
simulated_dataset %>%
  tabyl(q1_city, q11_rate_support)%>%
  adorn_totals("row", name = "Total")%>%
  adorn_percentages("row")%>%
  adorn_pct_formatting(digits=2)%>%
  adorn_ns()%>%
  adorn_title("combined",row_name = "City", col_name = "Rating")%>%
  rename(
    `Terrible` = `1`,
    `Poor` = `2`,
    `Fair` = `3`,
    `Good` = `4`,
    `Excellent` = `5`
  ) %>% 
  kable(align = 'ccccccc', caption = 'Helpfulness of Government Financial Support Programs on Restaurants') %>% 
  kable_styling(latex_options = "hold_position")
```

As shown in Figure \@ref(fig:q14), restaurants in Peterborough had applied more business loans than those in Brantford during the intervention. Since loans have to be paid back, businesses with low operating confidence were less likely to apply for loans. With increased business opportunities after reopening, restaurant owners were more likely to invest for faster recovery, and this corresponds with our previous findings of longer expected survival time (Figure \@ref(fig:q8)) and expected recovery time (Figure \@ref(fig:q10)) in Peterborough restaurants as mentioned before.

@@

In contrast, restaurants in Brantford generally received more grants than those in Peterborough(Figure \@ref(fig:q15)). Since grants are non-repayable funds and require certain conditions to apply, restaurants in Peterborough with better business conditions may not be eligible to apply for grants. And restaurants that suffered more in Brantford could receive more grants during the shutdown.

@@

To obtain an accurate estimate of the shutdown effect on restaurant businesses, average characteristics of restaurants in the reopening city and those in the shutdown city need to be identical (in theory) in the absence of the reopen intervention under current COVID outbreak. Since businesses in the food service industries are highly labor intensive [@cite_labor], the `total employment` in food service industries between selected cities is preferred to be as identical as possible. Besides, `population density` ( `Total Population`/ `Total Land Area in squared kilometers`) also has significant organizational life-cycle effects[@cite_density]. At different population density levels, competitions for food services varies, leading to different restaurants mortality rates even without the reopen intervention. As a result, total employment in food service and population density serve as the two primary restaurant operation characteristics for the selection of treatment city and control city.

@@

We used the employment by industry data and 2016 Census data available at Statistics Canada to identify cities with similar total employment in food service industry and population density. The above point graph(Figure \@ref(fig:selectplot)) indicated that Brantford and Peterborough have similar characteristics affecting restaurant operations. In addition, the coefficient of the graph $$\frac{Population\:Density}{Total \:Employment}=\frac{Total\:Population/Total\:Land\:Area}{Total\:Employment}$$, which could be interpreted as the number of people served by a restaurant employee per square kilometers, helps to identify the restaurant operating situation in designated regions. Similar coefficients between Brantford and Peterborough (0.306 vs 0.308) also suggest close restaurant characteristics. Besides, there is neither significant demographics nor land areas difference between these two cities, so we consider them as a valid comparison group and reopened Peterborough as the treatment group while maintaining Brantford shutdown as the control group.

@@

* The employment data at Statistics Canada uses the NAICS (North American Industry Classification System) standard and assigns employment services to Accommodation and Food Service category. So there is less employment in food services. We assume that the ratios between the accommodation employment and the food service employment between Peterborough and Brantford are the same.

* Since the distance between Brantford and Peterborough is more than 200 kilometers apart, residents in Brantford are unlikely to travel such long distance solely for a dine-in meal. So the treatment of reopening Peterborough would not affect the mealing behavior in Brantford. However, residents in shutdown regions near Peterborough might go for dine-in meals, potentially increasing sales in Peterborough restaurants. The actual reopening effect on restaurants could be overestimated due to possible increase in sales from nearby shutdown regions' residents that would not sustain once the reopen expanded to all Ontario regions.

* The mindful comparison between cities' similarities on employment, population density, etc. ensures the internal validity of our experiment so that we could accurately estimate the true impact of shutdown restrictions on restaurant. However, our shutdown effect evaluations may not be generalized to the entire population of interest, i.e., all restaurants in Ontario such as township regions or cosmopolitan city Toronto due to different levels of competitions, operating costs, etc.

@@

* Since results of the first survey before implementing our experiment suggest apparent similarities between restaurants in Peterborough and those in Brantford, we assume surveyed restaurants to have similar qualities. However, attributes of restaurants (capacity, employment, service type, etc) were not controlled in our analysis. In other words, we assume the response rate within the different restaurant subgroups to be the same. Variations in the actual response between restaurants in two cities are unknown and may impact the internal validity of results.

* Even though the size of our sampled population is somehow small (100 restaurants each in Peterborough and Brantford), it has already captured more than half of our frame population (There are no more than 200 restaurants in each city). So we believe our sampling population could represent the general characteristics of our frame population.

* The sample suffered from survivorship bias as we only surveyed restaurants that were still in operations. Restaurants that lost their businesses during the pandemic before October in 2020 could not be captured with our intervention started in October 2020, while these restaurants would be the most impacted by the COVID shutdowns. Another aspect of the survivorship bias involves the use of the same frame population for the first survey and the second survey. The three-month intervention would also affect business operating situations of restaurants. And the decreased response rate of the second survey might also imply the survivorship bias of our experiment.

* The sampling methods bore self-selection bias. From a psychological standpoint, respondents may have a stronger desire to respond the survey if they have a strong opinion to express about their situations. Because of increasing sales and growing operation confidence during the intervention, respondents in Peterborough restaurants might be more willing to participate to our survey, leading to the higher response rate in Peterborough.

@@

* Compared to the average survey response rate around `33%` [@cite_responserate], the response rates of our surveys are relatively higher (more than `50%` in the first survey and `40%` in the second survey). We credit our high response rates to efforts made in the following: (1) We spent a great amount of our budget to hire and train survey staffs to better approach our respondents. (2) We offered respondents chances to withdraw gift cards as incentives to participate our surveys. (3) We did not ask exact value of queation of our interest (e.g. Sales). Instead, we offer respondents several options to choose (e.g. `$0 ~ $10,000` ) so that respondents are more comfortable to participate in our survey.
